---
title: "Statistical Inference Project: ToothGrowth Analysis"
author: "Zikun wu"
date: "October 25, 2015"
output: html_document
---

# Overview
Conduct basic statistical inference for ToothGrowth dataset.

To see the R and R markdown files, visit my [github](https://github.com/qoolazy/Coursera-Statistical-Inference).

# Install and select the following library

```{r}
        library(dplyr)
        library(ggplot2)

```

# 1. Load the ToothGrowth data and perform some basic exploratory data analyses 

## Load dataset ToothGrowth
```{r}
            library(datasets)
            data(ToothGrowth)
```

## Some basic exploratory data analyses 
```{r}
            str(ToothGrowth)
            head(ToothGrowth)
            tail(ToothGrowth)
```

## Number of Rows and Columns.
```{r}

            dim(ToothGrowth)
 
```

## Sample Size n
```{r}

           length(ToothGrowth$len)

```

## Mean group by dose and by OJ & VC
```{r}

          aggregate(ToothGrowth$len,list(ToothGrowth$supp,ToothGrowth$dose),mean)

```

## Standard Deviation group by dose and by OJ & VC
```{r}

            aggregate(ToothGrowth$len,list(ToothGrowth$supp,ToothGrowth$dose),sd)

```

## Boxplot graph of the tooth length vs the does
```{r, echo=TRUE}
            ggplot(ToothGrowth, aes(x = factor(dose), y = len, fill = factor(dose)))+
                  geom_boxplot()+
                  facet_grid(.~supp)+
                  labs(title = "Tooth Length vs. Dose  by for OJ & VC",
                  x = "Doses", y = "Tooth Length")

```


# 2. Provide a basic summary of the data.

```{r}

            summary(ToothGrowth)

```
The dataset contains observations of the effect of Vitamin C on thooth growth in Guinea Pigs
It contains 60 observatins on 3 variables:

 * len - numeric tooth length
 * supp - a factor column for Supplement type (VC or OJ)
 * dose - numeric dose in milligrams

The response is the length of odontoblasts (teeth) in each of 10 guinea pigs at each of three dose levels of Vitamin C (0.5, 1, and 2 mg) with each of two delivery methods (orange juice or ascorbic acid).

## Table view of data. 
```{r}

            table(ToothGrowth$supp,ToothGrowth$dose)

```

As we can see there are 10 tests for each pair.

```{r}
ggplot(data = ToothGrowth, aes(x = as.factor(dose), y = len, fill = supp)) +
    geom_bar(stat = "identity") +
    facet_grid(. ~ supp) +
    xlab("Dose in miligrames") +
    ylab("Tooth length") +
    guides(fill = guide_legend(title = "Suppliment type"))
```

From the figure we can conclude that _there is a correlation between tooth growth size and dose levels of Vitamin C_, for both delivery methods. 

# 3. Use confidence intervals and/or hypothesis tests to compare tooth growth by supp and dose. 

From the box plot, OJ looks to be doing better with dose 0.5 and 1 effect on teeth growth than VC. To test this hypothesis by hold a the mean of OJ and VC does not cross zero.

## Dose 0.5
We are 95% confident that the limits of 1.719057 and 8.780943 actually do contain the difference between the two population means.Because those limts do not contain zero, this confodence interval suggests that it is very possible that the two population means are not equal.

```{r}

            ojdose0.5 <- ToothGrowth %>% filter(supp=="OJ" & dose=="0.5")
            vcdose0.5 <- ToothGrowth %>% filter(supp=="VC" & dose=="0.5")
            t.test(ojdose0.5$len,vcdose0.5$len)

```

## Dose 1
We are 95% confident that the limits of 2.802148 and 9.057852 actually do contain the difference between the two population means. Because those limts do not contain zero, this confodence interval suggests that it is very possible that the two population means are not equal.

```{r}

            ojdose1 <- ToothGrowth %>% filter(supp=="OJ" & dose=="1")
            vcdose1 <- ToothGrowth %>% filter(supp=="VC" & dose=="1")
            t.test(ojdose1$len,vcdose1$len)

```

## Dose 2
We are 95% confident that the limits of -3.79807 and 3.63807 actually do contain the difference between the two population means. However, because those limts do contain zero, this confodence interval suggests that it is very possible that the two population means are equal.

```{r}

            ojdose2 <- ToothGrowth %>% filter(supp=="OJ" & dose=="2")
            vcdose2 <- ToothGrowth %>% filter(supp=="VC" & dose=="2")
            t.test(ojdose2$len,vcdose2$len)

```

# 4. State your conclusions and the assumptions needed for your conclusions. 
We are 95% confident that dose 0.5 and dose 1 of OJ result in longer tooth length than dose 0.5 and dose 1 of VC. However, at the highest dose of 2, there is no statistically significant difference between the effects of OJ and VC. 
